---
title: "US Foreign Aid and Foreign Policy Preferences"
author: "Omar Cano"
description: "A GOV 50 Final Project on the impact on US  aid on  US and aid-receipient foriegn policy alignment"
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include = FALSE}
library(tidyverse)
library(tidyverse)
library(dplyr)
##install.packages("lubridate")
##install.packages("countrycode")
library(lubridate)
library(countrycode)

##load("~/Desktop/GOV 50/Projects/Final/final_project/data/un_votes.RData")


```


```{r ideal_data_prep, include = FALSE }
ideal <-  read.csv("data/IdealpointestimatesAll_Jul2023.csv")

## Adding year from session 
ideal <- ideal|> mutate( country_cowc = countrycode(ccode, origin = 'cown' , destination = 'cowc'),
                         country_cown = ccode,
                         year_match = session + 1945)
## W/O USA
ideal_wo_us <- ideal|> 
  group_by(country_cown, year_match)|>
  select(ccode, session, IdealPointAll, Countryname, year_match, country_cown, country_cowc)|>
  filter(!country_cown == 2)

## US Values only
ideal_onlyus <- ideal|>
  filter(country_cown == 2)|>
  select(session, IdealPointAll)|>
  mutate(idealpoint_us = IdealPointAll)|>
  select(!IdealPointAll)

## ideal now has US value
ideal<- ideal|> 
  left_join(ideal_onlyus, by = join_by(session))

ideal <- ideal|> 
  mutate(distance = idealpoint_us - IdealPointAll)|>
  mutate(distance_abs = abs(distance))

##Familiarizing Self w data
###ideal|> ggplot(mapping = aes(x = distance_abs))+geom_histogram()
##ideal|> ggplot(mapping = aes(x = distance))+geom_histogram()

## New objects for joining to ideal values of ideal point from previous years
minus_one <- ideal|> 
  group_by(ccode, session)|>
  select(ccode, session, distance_abs)|>
  mutate(session = session - 1)

minus_two <- ideal|> 
  group_by(ccode, session)|>
  select(ccode, session, distance_abs)|>
  mutate(session = session - 2)
 
plus_one <- ideal|> 
  group_by(ccode, session)|>
  select(ccode, session, distance_abs)|>
  mutate(session = session + 1 )

plus_two <- ideal|> 
  group_by(ccode, session)|>
  select(ccode, session, distance_abs)|>
  mutate(session = session + 2)


## Adding new columns of sideal values of ideal point from previous years two years

ideal <- ideal|> left_join(plus_one, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_plus_one"))


ideal <- ideal|> left_join(plus_two, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_plus_two"))

## Ordering and new titles for columns 
ideal<- ideal|> 
  mutate(distance_two_before = distance_abs_plus_two,
         distance_one_before = distance_abs_plus_one,
         distance_duplicate = distance_abs)

## Mutating for difference 
ideal<- ideal|> 
  mutate(change_one_before = distance_one_before - distance_duplicate,
         change_two_before = distance_two_before - distance_duplicate)


```



```{r aid_data_prep, include = FALSE}
## This is the aggregated data, can choose to use thes disaggregated later, if time 
aid <- read.csv("data/AidDataCoreDonorRecipientYear_ResearchRelease_Level1_v3.0.csv")

## Got correct number of row? Yes. 

##gives data from 1973 to 2013 
aid_us <- aid|>
  filter(donor == "United States")

##Change from  recepient name to COW character country code, add COW numeric code for matching w Ideal point data
aid_us <- aid_us|> mutate( country = countrycode(recipient, origin = 'country.name', destination = 'cowc'))
aid_us <- aid_us|> mutate( country_cown = countrycode(country, origin = 'cowc', destination = 'cown'))

aid_us_test<- aid_us |> filter(is.na(country_cown))

## Aid US, changing date, PLUS on year. In order to use previous year's aid commitments to the vote year
aid_us <- aid_us |> mutate(year_match = year + 1)


```

```{r testing_name_conversion, include = FALSE}
## Testing Names, there are no names that are ambiguously coded for COW C that show up in the UN voting data anyway, so there are no errors for COW N either 
##unique(aid_us_test$recipient)
##unique_names <- unique(ideal$Countryname) ## Possible problem  using ideal
##unique_names <- enframe(unique_names)


##cart|> filter(Countryname == "Serbia")


```


```{r joining_aid_ideal, include = FALSE}
##Checking that UN is unique? Good. 
##aid_us|>count(country_cown, year_match) |> filter(n > 1)

## Remove data for country_cown is N/A
aid_us <-  aid_us|>filter(!is.na(country_cown))

## Check that ideal point is unique? Good.
## ideal|>count(country_cown, year_match) |> filter(n > 1)

##Joined Aid and Ideal
ideal_aid <- ideal|> 
  left_join(aid_us, join_by(country_cown, year_match))

## adding column for zeros instead of NAs
ideal_aid<- ideal_aid |>
  mutate(committed = if_else( is.na(commitment_amount_usd_constant_sum), 0, commitment_amount_usd_constant_sum ))
         
## Adding millions
ideal_aid<- ideal_aid |>
  mutate(committed_millions = committed/ 1000000)



## Adding groups for aid determination
ideal_treat<- ideal_aid|> mutate(treatment_any =  if_else(committed_millions > 0, 1, 0),
                                 treat_some = if_else( (committed_millions > 0) & (committed_millions <= 1), 1, 0),
                                 treat_middle = if_else( (committed_millions > 1) & (committed_millions <= 100),1,0),
                                 treat_high = if_else(committed_millions > 100, 1, 0),
                                 treat_bil = if_else(committed_millions > 1000, 1, 0),
                                 treat_sig = if_else(committed_millions > 10, 1, 0),
                                 treatment_middle_up = if_else( (committed_millions > 1), 1 , 0),
                                 treatment_high_up = if_else(committed_millions > 100, 1, 0),
                                )

## Sub-setting to the data that we will actually use,  
ideal_treat_fil<- ideal_treat |>filter(session == 29:69)

```



# 1 Introduction
This paper will explore the relationship between a country's foreign-policy preferences and the amount of aid that country receives from the United States by looking at data that is derived from session 29-69 of the United Nations General Assembly. Ultimately the question that interests me is does US provision of aid make the recipients of aid more inclined to the support of the US-led liberal world order? 

The underlying foreign policy preferences of a state are a central concern in international relations literature-  “perhaps the most fundamental issue” in the discipline (Johnston 2005, 1040). It is of great interest to be able to predict which factors affect whether a state is likely to support one foreign policy over another. A state's preferences play a part in guiding how a state acts, but these preferences are not a "not directly observed and thus need to be inferred from observable behavior"(Dreher et al). Thus many turn to votes in the United Nations General Assembly as a place to observe state behavior and attempt to operationalize and quantify state preferences. UNGA votes are an attractive option primarily because "they are comparable and observable actions taken by many countries at set points in time"(Dreher et al). 

However, it is crucial to emphasize that there might be some important differences between a state's underlying preference and how a state chooses to vote in the UNGA. A  state's foreign policy preference might be just one among a number of factors that affect that state votes. For instance, while a country might vote the same way as the Unite States on all resolutions in the UNGA, other factors such as strategic calculation or coercion might be at play. That is to say, just because the voting patterns of a pair of countries align, it is not necessarily the case that that the preferences of these two countries aligns. For example, there is some evidence for US practice of using the IMF and World Bank to buy support from other member states for its position on votes in the United Nations General Council (reference Dreher and Strun 2012). While some anlaysis such as  (another is Dreher, Axel, P. Nunnenkamp, and R. Thiele. 2008 (‘‘Does US Aid Buy UN General Assembly Votes? A Disaggregated Analysis.’’ Public Choice 136 (1): 139-64.)) give a robust treatement of the quesiton regarding the effects of different types of aid, they are still interested in simply in the voting coincidence of the US and a recipient of its aid. However interesting this sort of analysis is, my paper aims to do something a bit different than a focus on vote-buying relationship between the the US and other countries but rather on the effect of US aid on underlying state-preference. 

My  research question is this: Do the state preferences of countries that receive US aid become more aligned with those of the US? My central hypothesis is that when the US commits to giving a country aid, it becomes more likely that the state preferences between these two countries align. A second hypothesis that I am interested in testing is whether or not the  recipients of greater levels of US aid are more likely to align, and align to a greater degree with the US than recipients of lower amounts of US aid. The null hypothesis is that variation in the foreign policy alignment between the the US and other countries is due to random chance and not to the amount of aid that a country receives from the US. 

WHY!!! Hyp? 
you need to present a plausible story why the hypothesis might be true. Often, this is in the form of a behaviorial or institutional explanation. As social scientists, we are not interested in idiosyncratic explanations; we want to understand systematic patterns and relationships!

Evidence for an increase in alignment between the US and another country would be to see a decrease in the distance between the foreign policy preference of these two countries. Thus evidence for my hypothesis would be to see that following the commitment of aid this distance would decrease, that is we would see a negative change after thes commitment of aid.  


##  2 Data and Methods 

In order to analyze this effect I use two pretty cool data sets, one for the dependent variable and one for the independent variable.

My dependent variable, `change_one_before`, is calculated from estimated measures of state preference provided by Dreher et. all 2008. This numeric estimate of state preference that Dreher et al provide is the result of a statistical model  that uses, in part, Eric Voeten's data on member-state votes in the United Nations General Assembly. This numeric measure of state-preference is given at a state/session level, that is it provides a single measure of state preference for a given country, in a given year. This ideal point data is more than just measure of state voting coincidence with the US-- this data set is one of the things that makes this paper interesting. This ideal point estimate is a ultimately, a measure to "consistently capture the position of states vis-a`-vis a US-led liberal order."(pg 431). Dreher et al explicitly emphasize the suitability of their ideal point estimates for analyzing the influence of aid, thus this is an interesting project that is taking this cool data from a statistical model and using it to attempt to learn about the dynamic that unfolds with US aid. 

There are a two steps to calculating my dependent variable `change_one_before`. First, I had to calculate a measure of the alignment of state preference between a given country and the US for a given session.  The variable `distance_abs` is a measure of the alignment of a given country  with the US in a given session, arrived at by subtracting the value of the a given countries ideal-point estimate for a particular session from the USA's ideal-point estimate for that same session, and finally taking the absolute value. Second, I use this `distance_abs` variable to calculate a measure of how a  countries preference alignment with the US has changed from the past session to the current session. Thus, my dependent variable `change_distance_one`, for a given country in session n is calculated by subtracting the `distance_abs`of session n from the `distance_abs` value for that country in session n-1.

Data for my independent variable, `treatment_any` comes from  AidData’s Core Research Release Version 3.0. This variable is a binary variable that is calculated from AidCores aggregated data (but also, they provide disaggreated data that is coded much more detail) on the amount of bilateral aid that a country reports that they committed to giving another country. This includes aid that was reported from various sources and is given in constant ($USD for 2011) dollars, and in million to allow for time-series analysis. This raw data is reported under the variable `commitment_millions`. This data is also reported at a recipient-year level, meaning that it provides the amount that the United State's committed to a given state in a given year. I follow the practice of assuming that if a country didn't report aid to  a country in a given year then the commitment of aid for that year was in fact zero. I also take a commitment to providing aid as a sufficient variable for analysis though there might be some discrepancy for aid that actually is paid out and aid that is committed, and thus some room for further analysis of aid actually delivered as the treatment condition.


```{r dep_plot, include = FALSE}
change_one_sd <- sd(ideal_treat_fil$change_one_before, na.rm = TRUE)
mean_dep<- mean(ideal_treat_fil$change_one_before, na.rm = TRUE)


dep_plot<- ideal_treat_fil|>
  ggplot(mapping = aes(x = change_one_before))+
  geom_histogram(binwidth = .07)+
  labs(title = "Figure One: Distribution of Dependent Variable",
       subtitle = "All included",
       x = "Change in alignment with USA from Previous Year, change_one_before",
       y = "Count")+
  geom_vline( xintercept = mean_dep)+
  geom_vline( xintercept = change_one_sd*2)+
  geom_vline( xintercept = change_one_sd*(-2) )
```
  

If we look at all the distribution of all `change_one_before` for all session, it appears that the variable follows a normally distribution, centered at about `mean_dep`, with a standard deviation of `change_one_sd` . 

 display dep_plot eventually

My research design employs a differences-in-differences approach. I treat the the commitment of aid as the treatment condition and want to investigate its effect on the alignment of state preferences. My independent variable is `treatment_any` and my depended variable is `change_one_before`. Both of my variables are determined for country-session pairs.

My independent variable is `treatment_any` is a binary variable describing whether or not a country was the recipient of any non-zero amount of US foreign aid in the preceding year. The control is no aid commitment reported as `0`, and receiving aid is the treatment reported as `1`.  Said differently, I lag the treatment variable to make it so that for a given session n, a  country is counted as being in the control group if and only if it received a commitment for aid in the the year before that session. Lagging the time of the of the commitment by one year is to provide the opportunity to observe an effect of treatment, whose effect's we might miss if we weren't to lag treatment. 

My dependent variable is`change_one_before` is a numeric variable describing the change in foreign policy alignment with the United States from the previous year to the current year. For a given country in session n we can calculate the difference in how that countries alignment with the United States has changed from the previous session, n-1. There is room for some confusion that should be preempted, an increase in alignment would be indicated by a negative value for `change_one_before`, because this would indicate a decrease in the distance between the foreign policy preferences of the US and a given country. 

To calculate the difference in differences we can then first take the average change among all countries in the control group and subtract it from the average change of all countries in the treatment group. This gives us an estimated average treatment effect. We can also run a regression of the values 

I work under the assumption that in a given session all countries will experience, on average parallel trends through time. Thus, if we observe a decrease in the distance from the USA for countries that are in the treatment group that is relatively higher than the decrease that we observe for those in the treatment, this would be evidence for my hypothesis. That is a statistically significant negative ATE would be evidence for my hypothesis. 

## 3 Results
plot of main analysis + regression output + 2-3 paragraphs of description and interpretation of the plots and regression (including interpreting the main coefficient of interest and describing if it is statistically significant and if we should interpret it causally). This section could be longer if you choose to include additional analyses. (8pts)

In summary my findings through the proposed difference in difference approach are rather mixed and inconclusive. While the regression that I arrive it does predict that on average the commitment of any amount of bilateral aid from the US to another country is correlated with a foreign-policy preference convergence between the US and that country, this prediction is not statistically significant and care should be taken to not construe any of my suggestions as casual inferences. 

My first and primary analysis used a difference in differences approach in order to calculate the average treatment effect. In order to perform this anlysis I run a regression analysis  on the treatment as a categorical predictor of the `change_one_before`. This regression yields two coefficients, one is an intercept which represents the models prediction for the baseline group, that is the control group. The value of the intercept is `intercept` indicating that on average countries in the control were slighly more likely to increase their distance from the US. However it is imporant to keep in mind that For countries in the control this model is rather uninformative, this intercept coefficient is not statistically significant, and in fact is highly arbitrary. For countries in the control this model is rather uninformative.  

For for countries in the treatment group the model provides a more clear prediction. The main coefficient of interest is equivalent to the average treatment effect that is reported in table __.  This coefficient is `main_coef`. It is negative, thus predicting that for a country that has received any amount of aid we should expect to see the distance between the ideal points of those two counties to decrease by approximately `-1*main_coef`. Recall, this decrease is an indication of increaed alignment. However, with a p-value of `main_coef_p`, this coeficient is also not statistically significant. Thus, with this approach I cannot reject the null hypothesis of any observed difference being due to pure chance. However, in the following section I perform another test, but it too is not statistically significant and cannot be used to derive any causal inferences. 

Need to make each of the following
table_formatted 
table_sum_ate
fit_ate_formated
avg_ate

This regression analysis reflects the same underlying approach as the difference in differences that can be calculated as in the table below. In this chart, I display the change in the distance between the US and other countries from the previous session to the current session. This table lists the data on which I ran the above  regression. Table __ summarizes the insights from this table by taking the averages across the groups, across sessions and fining the difference to estimate the average treatment effect. 

Whatever the results of my anlysis could have been, there is one important thing to note about the use of my dependent variable that limits any causal interpretation that we might derives from its use. For  all the virtues of the ideal-point estimates that Dreher et al provide, when used as I have used them, to compare the distance or alignment of two countries, these estimates of state alignment of state preference can  give no directional relationship btween countries. That is to say,  we cannot conclude with certainty, even from a statistically significant result, that the provision of US aid is causing the recipient of aid  to move closer to the foreign policy preference of the US. It just as well be that the US is moving closer in foreign policy preference to the countries to which it provides aid.



```{r reg_main, include = FALSE}

## Creating table for ATE in a given session
ate_table <-ideal_treat|> 
  group_by(session, treatment_any)|>
  summarize(mean_change_one = mean(change_one_before, na.rm = TRUE))|>
  pivot_wider(names_from = treatment_any,
              values_from = mean_change_one)|>
  mutate(session_ate = `1` - `0`)|>
  filter(!is.na(session_ate))

ate_table_names<- ate_table[, c(1, 3, 2, 4)]
colnames(ate_table_names) <- c("Session", "Treatment", "Control", "Session ATE")

## Gives table for ATE in a given session 
table_formatted<- knitr::kable(ate_table_names)


avg_ate <- mean(ate_table_names$`Session ATE`)


## Making ATE table longer, for visualization
ate_table_longer<- ate_table|> 
  select(!session_ate)|>
  pivot_longer(
    cols = `0`:`1`,
    names_to = "treatment",
    values_to = "change_one_before"
  )|>
  mutate(treatment = as.numeric(treatment))

## This is a geom_point showing treatment and control means across all sessions that we can compare with this data
ate_table_longer_point<- 
  ate_table_longer|> 
  ggplot(mapping = aes(x = session,
                       y = change_one_before))+
  geom_point(mapping = aes(color = treatment))

##Regression Plot for the Session ATEs.
ate_table_longer_num <-ate_table_longer|> 
  mutate(treatment = as.numeric(treatment))

ate_table_longer_num|> 
  ggplot(mapping = aes(x = treatment,
                       y = change_one_before))+
  geom_point()+
  geom_smooth(method = "lm", 
              formula = y ~ x,
              se = FALSE
              )

## Regression 
fit_ate_table_any<- lm(change_one_before ~ treatment, data = ate_table_longer)
sum_fit_ate <- summary(fit_ate_table_any)$coefficients

intercept<- sum_fit_ate[1,1]
int_p<- sum_fit_ate[1,4]

main_coef<- sum_fit_ate[2,1]
main_ceof_p<- sum_fit_ate[2,4]

fit_ate_formated <- modelsummary::modelsummary(fit_ate_table_any,
statistic = c("s.e. = {std.error}",
"p = {p.value}"),
gof_map = c("nobs", "r.squared", "adj.r.squared"))


```









## 5 Secondary Analysis
Because of the failure of my primary analysis to produce a statistically significant result, I performed this second analysis, in part to speak to the second hypothesis that I posed. I hypothesizes that higher amounts of aid could be correlated with an increased probablity of alignment with the US. Should we focus on analyzing the relationships between state preference and aid commitments, does a clear statistically significant pattern emerge for countries that are committed higher amounts of aid.

To test for this hypothesis I used I simply too pooled all observation of changes in the alignment with USA. Whereas before I was comparing a difference in difference among treatment and control for a given session, I now attempted to compare the average ATE by simply pooling all data on the changes in alignment with the US with no respect to session. Thus I get a distribution for all observed changes that looks like this `second_dist`.

I ran two regression models on this data. In the first `fit_treat_all_formated` I use the same treatment variable from my main analysis as a categorical predictor of the `change_one_before`. This regression yields two coefficients, whose values and corresponding p-values are summarized in__  Again we see that for countries in the control this model is rather uninformative, as thes intercept coefficient is not statistically significant. The slope coefficient is the coefficient of interest and again is negative, thus predicting that for a country that has received any amount of aid we should expect to see the distance between the ideal points of those two counties to decrease by approximately `-1*main_coef`. Recall, this decrease is an indication of increased alignment in foreign policy. However, with a p-value of `main_coef_p`, this coefficient is also not statistically significant.

The second regression `fit_treat_all_multi_formated` I run includes a two categorical variables as predictors, this time in attempt to see if there is any difference between the amount of aid that is provided. In this regression I now include two different categorical predictors. The first being `treatment_middle` for when a country receives a commitment between 1 and 100 million, and `treatment_high` for when a country receives a commitment greater than 100 million dollars. This regression yields a couple things of note, first that each of the coefficients of interest are negative, in line with my expectations about the decrease in distance, or said differently the increase in foreign-policy alignment. This might be construed of evidence for my first hypothesis-- that there is some decrease in distance when the US commit aid. The coefficients for both of these predictors are bigger in magnitude than that in my firt regresssion analysis. It might also be construed as support for my second hypothesis, as it was seen that perhaps we just needed to label as treatment only higher amount of committed aid in order to see a more pronounced correlation with a foreign policy preference alignment. The coefficient on the `treatement_middle` predictor is particularly of note. For a country that receives between  1 and 100 million dollars this model predicts an average decrease in distance of 0.09632 points, and this prediction is the closest my analysis has come to statistical significance, with a p-value of 0.190. 



```{r dis_chart_second, include = FALSE}
##Distribution Chart for all observed country changes in the period of interest
second_dist<- ideal_treat_fil|> 
  mutate(treatment_any = as.factor(treatment_any))|>
  ggplot(mapping = aes(x = change_one_before,
                        fill = treatment_any,
                        color =  treatment_any))+
  geom_histogram()

## Using multiple variables, multiple levels of aid
fit_treat_all <- lm (change_one_before ~ treatment_any, data = ideal_treat_fil)

fit_treat_all_multi <- lm (change_one_before ~ treat_middle + treat_high, data = ideal_treat_fil)

## Formatted 
fit_treat_all_formated <- modelsummary::modelsummary(fit_treat_all,
statistic = c("s.e. = {std.error}",
"p = {p.value}"),
gof_map = c("nobs", "r.squared", "adj.r.squared"))

fit_treat_all_multi_formated<- modelsummary::modelsummary(fit_treat_all_multi,
statistic = c("s.e. = {std.error}",
"p = {p.value}"),
gof_map = c("nobs", "r.squared", "adj.r.squared"))
```



## 6 Conclusion 
a brief (one paragraph) concluding section that summarizes your results, assesses the extent to which you find support for your hypothesis, describes limitations of your analysis and threats to inference, and states how your analysis could be improved (e.g., improved data that would be useful to collect).1 paragraph (i) summarizing results and assessing the extent to which you find support for your hypothesis; (ii) describing limitations of the analysis and threats to inference (missing data, confounding, etc), and stating how you could improve your analysis if you had more time/money. (2pts)

In conclusion, this paper was an exploration of the data that hoped to illuminate the relationship between foreign-policy preferences and US provision of foreign aid. A difference in differences approach treated the the commitment to providing any amount of aid as a treatment condition and analyzed the effect of this condition on the foreign policy alignment of a country with the US. My finding were ultimately mixed and inconclusive. While the regression that I arrive it does predict that on average the commitment of any amount of bilateral aid from the US to another country is correlated with a foreign-policy preference convergence between the US and that country, this prediction is not statistically significant and care should be taken to not construe even a statistically significant result as casual.A second analysis that pooled all available data and treated it as a cross-section analysis of the effect of aid, provided simiarly suggestive but inconclusive results that were no statistically significant. Thus, evidence for my hypothesis is weak, and I can not rule with any interesting level of probability that the result I observed are due to random chance. 

While this paper was unable to demonstrate statistically significant evidence for the hypothesis it shows some promise as a project for further exploration. There are ample opportunities to improve on this anlysis. Among them are threats to the inference, namely confounding varibales that are not being controlled for in any of the analysis that I performed. Another is that while the data set that was used for my independent variable is detailed, it is still the limiting factor in the range of sessions that I was able to analyze. This missing data kept me from including in my analysis the exiting state-preference data on the first 28 sessions of the United Nations. Similarly one important thing to note about the use of my dependent variable is that for all the virtures of the ideal-point estimates that Dreher et al provide, when used as I have used them, to compare the distance or alignment of two countries, their estimates give no directional relationship. That is to say,  we cannot conclude with certainty, even from a statistically significant result, that the provision of US aid is causing the recipient of aid  to move closer to the foreign policy preference of the US. The data and approach that I used for my dependent variable is mute on the point of just how two countries arrived at the new distance of their preference - it does not rule out any combination of both donor and recipient changing their preferences.If I were to have more time to perform this analysis, the most fruitful addition would be to identify and include data on possible confounders and to include this in a multi-variable regression. It is possible that the inclusion of such variables in my regression anlysis might just have produce a substantively similar difference in means as I found in this paper, but one that would display a statistical significance. 

## Works Cited



