---
title: "Gov 50 Final Project"
author: "Omar Cano"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

Note to self.... Finish a simple analysis.. 
If there is time later, then come back and edit or add to it. By tonight, a draft that could be submitted. 
 

## 1 Introduction, Reasearch Question, Hypthesis

I am interested in exploring data related to which state-level attributes contribute to member-state voting patterns in international organizations. International relations is broadly my area of interest, and I am interested in looking at when there is higher or lower voting coincidence with the US, whether under either democratic or republican leadership of the United States.

The question is really this, what is the difference between voting  alignment with the US stance under republican or democratic administrations? 

My hypothesis is that under democratic presidents we will see an increase in voting alignment than under republican administrations. I hypothesize this because ____.

The null hypothesis is that there is no significant difference between the measure of US-voting coincidence and the administration under which that coincidence is calculated.

## 2 Data Section
I will use Eric Voeten's member-state votes in the United Nations General Assembly. I will explore differences between UNGA voting in the periods of Carter to Obama.

My variable of interest is the measure of US alignment in the UNGA.
My explanatory variable is whether or not there is a Republican or Democrat in the White House. 

I will cover the years from the Cater Administration until the Obama Administration. 

This dataset needs to be altered in order to account for which administration votes occurred in. I also need to assign weights in order to calculate voting coincidence.

```{r loadata, echo = FALSE}
library(tidyverse)
library(dplyr)
install.packages("lubridate")
library(lubridate)

load("~/Desktop/GOV 50/Projects/Final/final_project/data/un_votes.RData") ## 

ideal <-  read.csv("data/IdealpointestimatesAll_Jul2023.csv")
lead <- read_dta("data/WhyLeadersFightLEADDataset_updated.dta")
presidents <- read.csv("data/presidents.csv")

presidents <- presidents|>
  select(Presidency, President, Took.office, Left.office, Party)

View(ideal)
View(completeVotes)
View(presidents)
## Create the presidents tibble for dates. 

completeVotes|> summary()

```

### 2.1 Preparing Data
```{r}
## Editing Dates on President's to Match UNGA Votes Format 
presidents <- presidents |> 
  mutate(left_office = as.Date(Left.office, format = "%d/%m/%Y"),
         took_office = as.Date(Took.office, format = "%d/%m/%Y"))

## Vote data to correct date format
completeVotes2 <- completeVotes|> mutate(
  date = as.Date(date, format = "%Y-%m-%d"))

View(completeVotes2)

##Subset UN Data, want only carter to Obama Votes
carter_obama <- completeVotes2|>
  filter( date > "1977-01-20" & date < "2017-01-20") ## needed to use just greater or lesser than

View(carter_obama)

##creating vector of democrat presidents 
democratic_indiators <- c("1", "4", "6")

##Assigning Presidency, Rep or Dem, and also president identifier 1 for carter, 2 for Reagan etc 
cart <- carter_obama|>
    mutate(pres_indicator = case_when( (date > "1977-01-20" & date < "1981-01-20") ~ "1",
                                       (date > "1981-01-20" & date < "1989-01-20") ~ "2",
                                       (date > "1989-01-20" & date < "1993-01-20") ~ "3",
                                       (date > "1993-01-20" & date < "2001-01-20") ~ "4",
                                       (date > "2001-01-20" & date < "2009-01-20") ~ "5",
                                       (date > "2009-01-20" & date < "2017-01-20") ~ "6")
           )|>
  mutate(democrat = if_else(pres_indicator %in% democratic_indiators, "1", "0"))|>
  mutate(party = if_else(democrat == 1, "Democrat", "Republican"))|>
  mutate(president = case_when(pres_indicator == 1 ~ "Jimmy Carter",
                               pres_indicator == 2 ~ "Ronald Reagan",
                               pres_indicator == 3 ~ "George H. W. Bush",
                               pres_indicator == 4 ~ "Bill Clinton",
                               pres_indicator == 5 ~ "George W. Bush",
                               pres_indicator == 6 ~ "Barack Obama")
  )

## Tibble of USA votes
us_votes <- cart|> 
  filter(Country == "USA")|>
  select(rcid, vote)|>
  mutate(us_vote = vote)|>
  select(!vote)

View(us_votes)

##Merge with all other votes, to indicate US vote in each row
cart <- cart |>
  left_join(us_votes, by = join_by(rcid))


## Adding Indicator of Voting With US or Not 
cart <- cart|> 
  mutate(coin_us = case_when((vote == 1 & us_vote == 1) | (vote ==3 & us_vote ==3) ~"Both Affirm/Deny",
                              vote %in% c("2","8") & us_vote %in% c("2","8") ~"Both Abstain/Absent",
                              (vote == 1 & us_vote ==2) | (vote == 2 & us_vote == 1) ~"Contra",
                             vote %in% c("2", "8") & us_vote %in% c("1", "2") ~ "Mixed"))

## Assigning Weights 
cart <- cart|> 
  mutate(weight_coin_us = case_when((vote == 1 & us_vote == 1) | (vote ==3 & us_vote ==3) ~ 1,
                              vote %in% c("2","8") & us_vote %in% c("2","8") ~ .5,
                              (vote == 1 & us_vote ==2) | (vote == 2 & us_vote == 1) ~0,
                             vote %in% c("2", "8") & us_vote %in% c("1", "2") ~ .5))

View(cart)

## Getting ATE per Country 
per_country_party <- cart|> 
  group_by(Country, party)|> 
  summarize(mean_country_coin = mean(weight_coin_us, na.rm = TRUE))|>
  pivot_wider(names_from = party,
              values_from = mean_country_coin)|>
  mutate( ate = `Democrat` - `Republican`)
  
View(per_country_party)


```

### 2.3 Dependent and Independent Variable 





## Running Calculations of Coincidence Means

```{r}
##Counting Number of Votes Total by Party
cart |> group_by(party)|>
  summarize(total = length(unique(rcid)))

##Mean Coincidence by Party
cart|> group_by(party)|>
  summarize(mean_coin = mean(weight_coin_us, na.rm = TRUE))

##Mean Coincidence per President
cart|> group_by(president)|>
summarize(mean_coin = mean(weight_coin_us, na.rm = TRUE))

##Maybe remove the israel votes? Maybe look at 
##Maybe look at who votes in line with the US most often?

cart|> group_by(Country, party)|>
  summarize(mean_coin_country = mean(weight_coin_us, na.rm = TRUE))|>
  pivot_wider(names_from = party, 
              values_from = mean_coin_country)|>
  mutate(ate = `Democrat` - `Republican`)|>
  ggplot(mapping = aes( y = ate))+
  geom_histogram()
```



```{r}
completeVotes |> filter (importantvote ==1, year >= 2018)
```








## Scrap Writing



I want to explore the extent to which there is state agreement on certain votes? 
The effect of something on voting coincidence in a particular time period and in a particular topic? 

The effect of political turn-over in the UNGA voting coincidence?
The role of leaders?
The role of some country-level feature?
 
 
Do small states agree more, do large states disagree more?
Democratic or Republican presidents at the UN?
Are there any statistical differences?



Which subset of states? 
Which votes, how to measure voting coincidence. Mutate,and create new variable that states coindidence or no coindidence. There are 

Domestic Turn-over, or... international factors.. 

 
What are valid and easy paper topics?
I just need either an easy GOV 50 paper, or an easy overlap between the GOV 50 paper and the UN paper. 

Do work on the UN paper first. Maybe whether or not there is an overlap. 


Can also just do a single-time analysis. 
Need something to test about the countries. '



```{r}
# Testing trump 
votes_trump<- completeVotes|>
  filter(year >= 2017)|>
  group_by(rcid)|> 
  summarize(mean = mean(vote ==1) )

votes_trump

View(completeVotes)

unique(completeVotes$year)
votes_trump|> summarize(mean = mean(mean))


votes_obama <- completeVotes|>
  filter(year < 2017 & year>2008)|>
  group_by(rcid)|> 
  summarize(mean = mean(vote ==1) )
votes_obama|> summarize (mean = mean(mean))


  group_by(rcid)|>
  summarize(mean_yes = mean(vote == 1))|>
  ggplot( mapping = aes( x =mean_yes))+
  geom_histogram(binwidth = .05)

## From LEAD data
lead_trump<- lead |> filter(year >= 2019)
lead
unique(lead_post1975$idacr)

```
