---
title: "Gov 50 Final Project"
author: "Omar Cano"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

```{r}
library(tidyverse)
```

## 1 Introduction: introduce the research question and hypothesis and briefly describe why it is interesting (1-2 paragraphs)
This paper will explore the relationship between a county's votes in an international organization and the amount of aid that country receives from the United States. There has been some evidence for the relationship of using the IMF and world bank to buy votes (reference Dreher and Strun 2012). What my paper is exploring is more a measure of state preferences as measured from the votes in the United Nations).  That is I will look at how on average astate's preferences relate to the amount of aid that it receives from the United States. sLooking not only at how does a country buy votes, but how much a state might come to align with the US through receiving aid. There is also this paper? Does it use ideal points? (Dreher, Axel, P. Nunnenkamp, and R. Thiele. 2008. ‘‘Does US Aid Buy UN General Assembly Votes? A Disaggregated Analysis.’’ Public Choice 136 (1): 139-64.) Dreher et al explicitly state the particular suitability of their ideal point estimates for analyzing the influence of aid, thus this is an interesting project that is taking this cool data from a statistical model and using it to attempt to learn about a particularly interesting and well-suited question. 
 
Generally explaining the voting patterns of member states in the international organizations is an object of interest to the international relations theorist. I will look at how well a state's level of aid from the USA can explain that state's level of voting coincidence with the US in the United Nations General Assembly. While voting coincidence is one things it is really state preference that I am interested in.

My  research question is:  Are the recipients of US aid more inclined to US-aligned voting in the UNGA? How might this aid change their state preference.

My central hypothesis is that through time, recipients of US aid vote come to vote more frequently in line with the US.T his is interesting, as a hypothesis because ___.

A second hypothesis is that recipients of greater levels of US aid are more likely to vote in-line with the US than recepeints of lower amounts of US aid. 

The null hypothesis is that there any difference between the measure of US-voting coincidence and the amount of aid that a country receives is due to pure chance. 

## 2 Data Section ---------- Briefly describes the data source, describes how the key dependent and independent variables are measured (e.g., a survey, statistical model, or expert coding), and also produces a plot that summarizes the dependent variable) (2-3 paragraphs)

This paper relies on two data sources, one for the dependent variable and one for the independent variable.

My dependent variable, `distance`, is calculated from data of estimated measures of state preference as provided by Dreher et. all 2008. This numeric estimate of state preference that Dreher et al provide is the result of a statistical model derived from Eric Voeten's data on member-state votes in the United Nations General Assembly. This numeric measure of state-preference is given at a state/session level, that is it provides a single measure of state preference as calculated from United Nations General Assembly Voting Data for a given country, in a given year. More specifically this estimate is a measure of "The resulting estimates consistently capture the position of states vis-a`-vis a US-led liberal order."(pg 431)

My dependent variable `distance` is a measure for how aligned two state's preferences are.The alignment of state preference for any pair of countries in a given session can be estimated by taking each country's ideal-point estimate in the same session,  subtracting them - that is finding the "distance" between these two points,and taking the absolute value. Thus, my dependent variable `distance` is a measure of the alignment of a given country  with the US in a given session, arrived at by subtracting the value of the a given countries ideal-point estimate for a particular session from the USA's ideal-point estimate for that same session. The value for the US in any given year is zero, because the US displayed voting patterns identical to itself. A low value for `distance` indicates higher alignment in state preference.

Data for my independent variable, `aid_committment`, is from  AidData’s Core Research Release Version 3.0. This data provides aggregated data (but also, they provide disaggreated data that is coded much more detail) on the amount of bilateral aid that a country reports that they committed to giving another country. This includes aid that was reported in __ and is given in constant dollars ($USD for 2011) to  allow for time-series analysis. This variable is also reported at a recipient-year level, meaning that it provides the amount that was committed to a given state in a given year. In order to perform my analysis, I lag the aid year, in order to allow for that aid to take an effect.

Plot summarizing the DEPENDENT variable
```{r}

## Top 10 receivers of aid 
top10_rec<- ideal_aid|> group_by(country)|>
  summarize(total_aid = sum(commitment_amount_usd_constant_sum))|>
  top_n(10, total_aid)|> 
  pull(country)

## Plotting top 10 recievers of aid 
ideal_aid|>
  filter(country %in% top10_rec)|>
  ggplot(mapping = aes(x = year, y = distance))+
    geom_line(mapping = aes(color = country))

##Distributions of distances, all  through time
ideal_aid|>
  ggplot(mapping = aes(x = distance_abs))+
  geom_histogram()

##What causes a big drop in the data? Are there any changes that are bigger than others? 
ideal_aid |> group_by(ccode, session)|>
  summarize(mean_prev_years = sum(idealpin ))
  mutate(change_diff =  )

  
  
## finding prev years into rows,
minus_one <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, IdealPointAll)|>
  mutate(session_minus_one = session - 1)

minus_two <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, IdealPointAll)|>
  mutate(session_minus_two = session - 2)
 
plus_one <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, IdealPointAll)|>
  mutate(session_plus_one = session + 1 )

plus_two <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, IdealPointAll)|>
  mutate(session_plus_two = session + 2)

idea_aid <- ideal_aid|> mutate(session_minus_one = session - 1,
                   session_minus_two = session - 2,
                   session_plus_one = session + 1,
                   session_plus_two = session + 2
                   )

ideal_aid|> left_join(minus_one )

  
  ## W/O USA
ideal_wo_us <- ideal|> 
  group_by(country_cown, year_match)|>
  select(ccode, session, IdealPointAll, Countryname, year_match, country_cown, country_cowc)|>
  filter(!country_cown == 2)

## US Values only
ideal_onlyus <- ideal|>
  filter(country_cown == 2)|>
  select(session, IdealPointAll)|>
  mutate(idealpoint_us = IdealPointAll)|>
  select(!IdealPointAll)

## ideal now has US value
ideal<- ideal|> 
  left_join(ideal_onlyus, by = join_by(session))

ideal <- ideal|> 
  mutate(distance = idealpoint_us - IdealPointAll)|>
  mutate(distance_abs = abs(distance))

View(ideal_us)


```

 
## Results--------- plot of main analysis + regression output + 2-3 paragraphs of description and interpretation of the plots and regression (including interpreting the main coefficient of interest and describing if it is statistically significant and if we should interpret it causally). This section could be longer if you choose to include additional analyses. (8pts)

Plot of Main analysis
```{r}
ideal_aid|>
  ggplot(mapping = aes(x =  committed_millions,
                       y = distance_abs))+
  geom_point()

fit <- lm(distance_abs ~ committed_millions, data = ideal_aid)

summary(fit)

View(fit)
coef(fit)

View(ideal_aid)
```



Interpretation of Coeficients 
Stastical Significance?  

## Conclusion






```{r loadata, echo = FALSE}
library(tidyverse)
library(dplyr)
install.packages("lubridate")
install.packages("countrycode")
library(lubridate)
library(countrycode)

load("~/Desktop/GOV 50/Projects/Final/final_project/data/un_votes.RData") ## 

ideal <-  read.csv("data/IdealpointestimatesAll_Jul2023.csv")
lead <- read_dta("data/WhyLeadersFightLEADDataset_updated.dta")
presidents <- read.csv("data/presidents.csv")

presidents <- presidents|>
  select(Presidency, President, Took.office, Left.office, Party)

View(ideal)
View(completeVotes)
View(presidents)
## Create the presidents tibble for dates. 

completeVotes|> summary()

```

## Ideal Points  
```{r}
## Can use a countries ideal point estimate? 
## Can use a fraction of the voting decisions? Maybe use the ideal points first, if time think about the raw voting data
View(ideal)

## Adding year from session 
ideal <- ideal|> mutate( country_cowc = countrycode(ccode, origin = 'cown' , destination = 'cowc'),
                         country_cown = ccode,
                         year_match = session + 1945)
## W/O USA
ideal_wo_us <- ideal|> 
  group_by(country_cown, year_match)|>
  select(ccode, session, IdealPointAll, Countryname, year_match, country_cown, country_cowc)|>
  filter(!country_cown == 2)

## US Values only
ideal_onlyus <- ideal|>
  filter(country_cown == 2)|>
  select(session, IdealPointAll)|>
  mutate(idealpoint_us = IdealPointAll)|>
  select(!IdealPointAll)

## ideal now has US value
ideal<- ideal|> 
  left_join(ideal_onlyus, by = join_by(session))

ideal <- ideal|> 
  mutate(distance = idealpoint_us - IdealPointAll)|>
  mutate(distance_abs = abs(distance))

View(ideal_us)



## Familiarizing Self w data
ideal|> ggplot(mapping = aes(x = distance_abs))+
  geom_histogram()

ideal|> ggplot(mapping = aes(x = distance))+
  geom_histogram()

```


###Prepping Aid Data 
```{r}
## This is the aggregated data, can choose to use thes disaggregated later, if time 
aid <- read.csv("data/AidDataCoreDonorRecipientYear_ResearchRelease_Level1_v3.0.csv")
aid
## Got correct number of row? Yes. 

View(aid_us)

##gives data from 1973 to 2013 
aid_us <- aid|>
  filter(donor == "United States")

##Change from  recepient name to COW character country code, add COW numeric code for matching w Ideal point data
aid_us <- aid_us|> mutate( country = countrycode(recipient, origin = 'country.name', destination = 'cowc'))
aid_us <- aid_us|> mutate( country_cown = countrycode(country, origin = 'cowc', destination = 'cown'))

aid_us_test <- aid_us |> filter(is.na(country_cown))

## Testing Names, there are no names that are ambigously coded for COW C that show up in the UN voting data anyway, so there are no errors for COW N either 
unique(aid_us_test$recipient)
unique_names <- unique(cart$Countryname)
unique_names <- enframe(unique_names)
View(unique_names)
cart|> filter(Countryname == "Serbia")

## Aid US, changing date, PLUS on year. In order to use previous year's aid commitments to the vote year
aid_us <- aid_us |> mutate(year_match = year + 1)|> select(!year_lagged)

## Familiarizing Data.
aid_us|> ggplot(mapping = aes( x = year, y = commitment_amount_usd_constant_sum ))+
  geom_point()

country_amount <-aid_us |> 
   group_by(year_match)|> 
   summarize(num_countries  = length(unique(country_cown)),
             total_given = sum(commitment_amount_usd_constant_sum)) 
  


aid_us |>
  ggplot(mapping = aes(x = as.factor(year_match),
                               y = commitment_amount_usd_constant_sum) )+
  geom_boxplot()+
  scale_y_log10()
  
  
  
country_amount|>
  ggplot(mapping = aes(x = as.factor(year_match),
                       y = num_countries))+
  geom_line()


country_amount|>
  ggplot(mapping = aes(x = year_match,
                       y = total_given))+
  geom_line()

 
```

## Merging Aid and Idea 
```{r}
View(ideal)
View(aid_us)

##Checking that UN is unique? Good. 
aid_us|>
  count(country_cown, year_match) |>
  filter(n > 1)

## Remove data for country_cown is N/A
aid_us <-  aid_us|>filter(!is.na(country_cown))

## Check that ideal point is unique? Good.
ideal|>
  count(country_cown, year_match) |>
  filter(n > 1)

##Joined Aid and Ideal
ideal_aid <- ideal|> 
  left_join(aid_us, join_by(country_cown, year_match))
View(ideal_aid)

ideal_aid |>filter(year_match > 1990,
                   commitment_amount_usd_constant_sum >1000000)|>
  ggplot(mapping = aes(x = commitment_amount_usd_constant_sum, 
                                  y = distance))+
  geom_point()

## adding column for zeros instead of NAs
ideal_aid <- ideal_aid |>
  mutate(committed = if_else( is.na(commitment_amount_usd_constant_sum),0, commitment_amount_usd_constant_sum)
         
## Adding millions
ideal_aid <- ideal_aid |>
  mutate(committed_millions = committed/ 1000000)


## Adding group for aid determination.. is there some way of showing this? 
```



















## Scrap Writing
I want to explore the extent to which there is state agreement on certain votes? 
The effect of something on voting coincidence in a particular time period and in a particular topic? 

The effect of political turn-over in the UNGA voting coincidence?
The role of leaders?
The role of some country-level feature?
 
 
Do small states agree more, do large states disagree more?
Democratic or Republican presidents at the UN?
Are there any statistical differences?



Which subset of states? 
Which votes, how to measure voting coincidence. Mutate,and create new variable that states coindidence or no coindidence. There are 

Domestic Turn-over, or... international factors.. 

 
What are valid and easy paper topics?
I just need either an easy GOV 50 paper, or an easy overlap between the GOV 50 paper and the UN paper. 

Do work on the UN paper first. Maybe whether or not there is an overlap. 


Can also just do a single-time analysis. 
Need something to test about the countries. '

```{r}
##Counting Number of Votes Total by Party
cart |> group_by(party)|>
  summarize(total = length(unique(rcid)))

##Mean Coincidence by Party
cart|> group_by(party)|>
  summarize(mean_coin = mean(weight_coin_us, na.rm = TRUE))

##Mean Coincidence per President
cart|> group_by(president)|>
summarize(mean_coin = mean(weight_coin_us, na.rm = TRUE))

##Maybe remove the israel votes? Maybe look at 
##Maybe look at who votes in line with the US most often?

cart|> group_by(Country, party)|>
  summarize(mean_coin_country = mean(weight_coin_us, na.rm = TRUE))|>
  pivot_wider(names_from = party, 
              values_from = mean_coin_country)|>
  mutate(ate = `Democrat` - `Republican`)|>
  ggplot(mapping = aes( y = ate))+
  geom_histogram()
```

```{r}
# Testing trump 
votes_trump<- completeVotes|>
  filter(year >= 2017)|>
  group_by(rcid)|> 
  summarize(mean = mean(vote ==1) )

votes_trump

View(completeVotes)

unique(completeVotes$year)
votes_trump|> summarize(mean = mean(mean))


votes_obama <- completeVotes|>
  filter(year < 2017 & year>2008)|>
  group_by(rcid)|> 
  summarize(mean = mean(vote ==1) )
votes_obama|> summarize (mean = mean(mean))


  group_by(rcid)|>
  summarize(mean_yes = mean(vote == 1))|>
  ggplot( mapping = aes( x =mean_yes))+
  geom_histogram(binwidth = .05)

## From LEAD data
lead_trump<- lead |> filter(year >= 2019)
lead
unique(lead_post1975$idacr)

```



```{r}
## Editing Dates on President's to Match UNGA Votes Format 
presidents <- presidents |> 
  mutate(left_office = as.Date(Left.office, format = "%d/%m/%Y"),
         took_office = as.Date(Took.office, format = "%d/%m/%Y"))

## Vote data to correct date format
completeVotes2 <- completeVotes|> mutate(
  date = as.Date(date, format = "%Y-%m-%d"))

View(completeVotes2)

##Subset UN Data, want only carter to Obama Votes
carter_obama <- completeVotes2|>
  filter( date > "1977-01-20" & date < "2017-01-20") ## needed to use just greater or lesser than

View(carter_obama)

##creating vector of democrat presidents 
democratic_indiators <- c("1", "4", "6")

##Assigning Presidency, Rep or Dem, and also president identifier 1 for carter, 2 for Reagan etc 
cart <- carter_obama|>
    mutate(pres_indicator = case_when( (date > "1977-01-20" & date < "1981-01-20") ~ "1",
                                       (date > "1981-01-20" & date < "1989-01-20") ~ "2",
                                       (date > "1989-01-20" & date < "1993-01-20") ~ "3",
                                       (date > "1993-01-20" & date < "2001-01-20") ~ "4",
                                       (date > "2001-01-20" & date < "2009-01-20") ~ "5",
                                       (date > "2009-01-20" & date < "2017-01-20") ~ "6")
           )|>
  mutate(democrat = if_else(pres_indicator %in% democratic_indiators, "1", "0"))|>
  mutate(party = if_else(democrat == 1, "Democrat", "Republican"))|>
  mutate(president = case_when(pres_indicator == 1 ~ "Jimmy Carter",
                               pres_indicator == 2 ~ "Ronald Reagan",
                               pres_indicator == 3 ~ "George H. W. Bush",
                               pres_indicator == 4 ~ "Bill Clinton",
                               pres_indicator == 5 ~ "George W. Bush",
                               pres_indicator == 6 ~ "Barack Obama")
  )

## Tibble of USA votes
us_votes <- cart|> 
  filter(Country == "USA")|>
  select(rcid, vote)|>
  mutate(us_vote = vote)|>
  select(!vote)

View(us_votes)

##Merge with all other votes, to indicate US vote in each row
cart <- cart |>
  left_join(us_votes, by = join_by(rcid))


## Adding Indicator of Voting With US or Not 
cart <- cart|> 
  mutate(coin_us = case_when((vote == 1 & us_vote == 1) | (vote ==3 & us_vote ==3) ~"Both Affirm/Deny",
                              vote %in% c("2","8") & us_vote %in% c("2","8") ~"Both Abstain/Absent",
                              (vote == 1 & us_vote ==2) | (vote == 2 & us_vote == 1) ~"Contra",
                             vote %in% c("2", "8") & us_vote %in% c("1", "2") ~ "Mixed"))

## Assigning Weights 
cart <- cart|> 
  mutate(weight_coin_us = case_when((vote == 1 & us_vote == 1) | (vote ==3 & us_vote ==3) ~ 1,
                              vote %in% c("2","8") & us_vote %in% c("2","8") ~ .5,
                              (vote == 1 & us_vote ==2) | (vote == 2 & us_vote == 1) ~0,
                             vote %in% c("2", "8") & us_vote %in% c("1", "2") ~ .5))

View(cart)


## Getting ATE per Country 
per_country_party <- cart|> 
  group_by(Country, party)|> 
  summarize(mean_country_coin = mean(weight_coin_us, na.rm = TRUE))|>
  pivot_wider(names_from = party,
              values_from = mean_country_coin)|>
  mutate( ate = `Democrat` - `Republican`)
  
View(per_country_party)



```