---
title: "Gov 50 Final Project"
author: "Omar Cano"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to the influence of international organizations on the policy of their member-states. International relations is broadly my area of interest, and I am interested in looking perhaps at  how IOs are a force that can shape state's decision patterns in particular policy domains. 

## Reasearch Question
I have just recently changed my topic of interest for this final project. It is more in line with the porject thoughts outlined above. I aim to use Eric Voeten's member-state votes in the United Nations General Assembly.I have yet to refine my research question. 


Below was a research question that I have since rejected and moved on from.

I will use the "Correlates of War" Citation: Sarkees, Meredith Reid and Frank Wayman (2010). Resort to War: 1816 – 2007. Washington DC: CQ Press. I will use it to analyze the following research quetion: What was the relation between the Chinese foreign policy alignnment towards the Soviet Union and the balance of power? I hypothesize that the realignment was due to a decrease in the Chinese power and an increase in Russian power. My outcome variable will be a measure of similiarty between the core foreign policy objectives of China aligning with the core policy objectives of the Soviet Union. My explanatory variables will be a combination of military strength and trade relations. If we observe an increase of alignment alongside a increase in relative Soviet power, this would be evidence for my hypothesis. If we observe an increase in the level of alignment without a relative increase in power this would be evidence against my hypothesis. 



```{r loadata }
library(tidyverse)

load("~/Desktop/GOV 50/Projects/Final/final_project/data/un_votes.RData") ## 

```

## Data Visualization 
Though I have yet to narrow in on a question  below is a data visualization that is going to help oreint me to some of the relevant features of the data set. I have included...  

```{r}
 test <- completeVotes |>  
  group_by(session)|>
  summarize( votes = sum(hr))


test|>ggplot(mapping = aes( x = session,
                         y = votes))+
  geom_col()+
  labs(x = "Session Number",
       y = "Number of Votes on Human Rights")
  
  
  
               
```


