






 From data section 

```{r values, include = FALSE}
change_one_sd <- sd(ideal_treat_fil$change_one_before, na.rm = TRUE)
mean_dep<- mean(ideal_treat_fil$change_one_before, na.rm = TRUE)
```


```{r dep_plot, include = FALSE}

dep_plot<- ideal_treat_fil|>
  ggplot(mapping = aes(x = change_one_before))+
  geom_histogram(binwidth = .07)+
  labs(title = "Figure One: Distribution of Dependent Variable",
       subtitle = "All included",
       x = "Change in alignment with USA from Previous Year, change_one_before",
       y = "Count")+
  geom_vline( xintercept = mean_dep)+
  geom_vline( xintercept = change_one_sd*2)+
  geom_vline( xintercept = change_one_sd*(-2) )
  
```











dep_plot<- ideal_treat_fil|>
  ggplot(mapping = aes(x = change_one_before))+
  geom_histogram(color = "orange", 
                 fill = "orange",
                 binwidth = .07)+
  labs(title = "Figure One: Distribution of Dependent Variable",
       subtitle = "All included",
       x = "Change in alignment with USA from Previous Year, `change_one_before`",
       y = "Count")+
  geom_vline( xintercept = mean_dep)+
  geom_vline( xintercept = change_one_sd*2)+
  geom_vline( xintercept = change_one_sd*(-2) )



`r table_formatted`
`r table_sum_ate`
`r fit_ate_formated`
`r avg_ate`






## Second Reg
```{r second_reg, include = FALSE}

## Creating table for ATE in a given session
ate_table_two <-ideal_treat|> 
  group_by(session, treatment_middle_up)|>
  summarize(mean_change_one = mean(change_one_before, na.rm = TRUE))|>
  pivot_wider(names_from = treatment_middle_up,
              values_from = mean_change_one)|>
  mutate(session_ate = `1` - `0`)|>
  filter(!is.na(session_ate))

ate_table_names_two<- ate_table[, c(1, 3, 2, 4)]
colnames(ate_table_names) <- c("Session", "Treatment", "Control", "Session ATE")

## Gives table for ATE in a given session 
table_formatted_two<- knitr::kable(ate_table_names)


avg_ate_two <- mean(ate_table_names_two$`Session ATE`)


## Making ATE table longer, for visualization
ate_table_longer_two<- ate_table_two|> 
  select(!session_ate)|>
  pivot_longer(
    cols = `0`:`1`,
    names_to = "treatment_middle_up",
    values_to = "change_one_before"
  )|>
  mutate(treatment = as.numeric(treatment_middle_up))

## This is a geom_point showing treatment and control means across all sessions that we can compare with this data
ate_table_longer_point_two<- 
  ate_table_longer_two|> 
  ggplot(mapping = aes(x = session,
                       y = change_one_before))+
  geom_point(mapping = aes(color = treatment_middle_up))

##Regression Plot for the Session ATEs.
ate_table_longer_num_two<-ate_table_longer_two|> 
  mutate(treatment = as.numeric(treatment_middle_up))

ate_table_longer_num_two|> 
  ggplot(mapping = aes(x = treatment_middle_up,
                       y = change_one_before))+
  geom_point()+
  geom_smooth(method = "lm", 
              formula = y ~ x,
              se = FALSE
              )

## Regression 
fit_ate_table_any_two<- lm(change_one_before ~ treatment_middle_up, data = ate_table_longer_two)
sum_fit_ate <- summary(fit_ate_table_any_two)$coefficients

intercept<- sum_fit_ate[1,1]
int_p<- sum_fit_ate[1,4]

main_coef<- sum_fit_ate[2,1]
main_ceof_p<- sum_fit_ate[2,4]

fit_ate_formated_two<- modelsummary::modelsummary(fit_ate_table_any_two,
statistic = c("s.e. = {std.error}",
"p = {p.value}"),
gof_map = c("nobs", "r.squared", "adj.r.squared"))

```





































## Scrap Writing
I want to explore the extent to which there is state agreement on certain votes? 
The effect of something on voting coincidence in a particular time period and in a particular topic? 

The effect of political turn-over in the UNGA voting coincidence?
The role of leaders?
The role of some country-level feature?
 

Do small states agree more, do large states disagree more?
Democratic or Republican presidents at the UN?
Are there any statistical differences?



Which subset of states? 
Which votes, how to measure voting coincidence. Mutate,and create new variable that states coindidence or no coindidence. There are 

Domestic Turn-over, or... international factors.. 

 
What are valid and easy paper topics?
I just need either an easy GOV 50 paper, or an easy overlap between the GOV 50 paper and the UN paper. 

Do work on the UN paper first. Maybe whether or not there is an overlap. 


Can also just do a single-time analysis. 
Need something to test about the countries. '

```{r}
##Counting Number of Votes Total by Party
cart |> group_by(party)|>
  summarize(total = length(unique(rcid)))

##Mean Coincidence by Party
cart|> group_by(party)|>
  summarize(mean_coin = mean(weight_coin_us, na.rm = TRUE))

##Mean Coincidence per President
cart|> group_by(president)|>
summarize(mean_coin = mean(weight_coin_us, na.rm = TRUE))

##Maybe remove the israel votes? Maybe look at 
##Maybe look at who votes in line with the US most often?

cart|> group_by(Country, party)|>
  summarize(mean_coin_country = mean(weight_coin_us, na.rm = TRUE))|>
  pivot_wider(names_from = party, 
              values_from = mean_coin_country)|>
  mutate(ate = `Democrat` - `Republican`)|>
  ggplot(mapping = aes( y = ate))+
  geom_histogram()
```

```{r}
# Testing trump 
votes_trump<- completeVotes|>
  filter(year >= 2017)|>
  group_by(rcid)|> 
  summarize(mean = mean(vote ==1) )

votes_trump

View(completeVotes)

unique(completeVotes$year)
votes_trump|> summarize(mean = mean(mean))


votes_obama <- completeVotes|>
  filter(year < 2017 & year>2008)|>
  group_by(rcid)|> 
  summarize(mean = mean(vote ==1) )
votes_obama|> summarize (mean = mean(mean))


  group_by(rcid)|>
  summarize(mean_yes = mean(vote == 1))|>
  ggplot( mapping = aes( x =mean_yes))+
  geom_histogram(binwidth = .05)

## From LEAD data
lead_trump<- lead |> filter(year >= 2019)
lead
unique(lead_post1975$idacr)

```



```{r}
## Editing Dates on President's to Match UNGA Votes Format 
presidents <- presidents |> 
  mutate(left_office = as.Date(Left.office, format = "%d/%m/%Y"),
         took_office = as.Date(Took.office, format = "%d/%m/%Y"))

## Vote data to correct date format
completeVotes2 <- completeVotes|> mutate(
  date = as.Date(date, format = "%Y-%m-%d"))

View(completeVotes2)

##Subset UN Data, want only carter to Obama Votes
carter_obama <- completeVotes2|>
  filter( date > "1977-01-20" & date < "2017-01-20") ## needed to use just greater or lesser than

View(carter_obama)

##creating vector of democrat presidents 
democratic_indiators <- c("1", "4", "6")

##Assigning Presidency, Rep or Dem, and also president identifier 1 for carter, 2 for Reagan etc 
cart <- carter_obama|>
    mutate(pres_indicator = case_when( (date > "1977-01-20" & date < "1981-01-20") ~ "1",
                                       (date > "1981-01-20" & date < "1989-01-20") ~ "2",
                                       (date > "1989-01-20" & date < "1993-01-20") ~ "3",
                                       (date > "1993-01-20" & date < "2001-01-20") ~ "4",
                                       (date > "2001-01-20" & date < "2009-01-20") ~ "5",
                                       (date > "2009-01-20" & date < "2017-01-20") ~ "6")
           )|>
  mutate(democrat = if_else(pres_indicator %in% democratic_indiators, "1", "0"))|>
  mutate(party = if_else(democrat == 1, "Democrat", "Republican"))|>
  mutate(president = case_when(pres_indicator == 1 ~ "Jimmy Carter",
                               pres_indicator == 2 ~ "Ronald Reagan",
                               pres_indicator == 3 ~ "George H. W. Bush",
                               pres_indicator == 4 ~ "Bill Clinton",
                               pres_indicator == 5 ~ "George W. Bush",
                               pres_indicator == 6 ~ "Barack Obama")
  )

## Tibble of USA votes
us_votes <- cart|> 
  filter(Country == "USA")|>
  select(rcid, vote)|>
  mutate(us_vote = vote)|>
  select(!vote)

View(us_votes)

##Merge with all other votes, to indicate US vote in each row
cart <- cart |>
  left_join(us_votes, by = join_by(rcid))


## Adding Indicator of Voting With US or Not 
cart <- cart|> 
  mutate(coin_us = case_when((vote == 1 & us_vote == 1) | (vote ==3 & us_vote ==3) ~"Both Affirm/Deny",
                              vote %in% c("2","8") & us_vote %in% c("2","8") ~"Both Abstain/Absent",
                              (vote == 1 & us_vote ==2) | (vote == 2 & us_vote == 1) ~"Contra",
                             vote %in% c("2", "8") & us_vote %in% c("1", "2") ~ "Mixed"))

## Assigning Weights 
cart <- cart|> 
  mutate(weight_coin_us = case_when((vote == 1 & us_vote == 1) | (vote ==3 & us_vote ==3) ~ 1,
                              vote %in% c("2","8") & us_vote %in% c("2","8") ~ .5,
                              (vote == 1 & us_vote ==2) | (vote == 2 & us_vote == 1) ~0,
                             vote %in% c("2", "8") & us_vote %in% c("1", "2") ~ .5))

View(cart)


## Getting ATE per Country 
per_country_party <- cart|> 
  group_by(Country, party)|> 
  summarize(mean_country_coin = mean(weight_coin_us, na.rm = TRUE))|>
  pivot_wider(names_from = party,
              values_from = mean_country_coin)|>
  mutate( ate = `Democrat` - `Republican`)
  
View(per_country_party)





```






```{r}

## Top 10 receivers of aid 
top10_rec<- ideal_aid|> group_by(country)|>
  summarize(total_aid = sum(commitment_amount_usd_constant_sum))|>
  top_n(10, total_aid)|> 
  pull(country)

## Plotting top 10 recievers of aid 
ideal_aid|>
  filter(country %in% top10_rec)|>
  ggplot(mapping = aes(x = year, y = distance))+
    geom_line(mapping = aes(color = country))

##Distributions of distances, all  through time
ideal_aid|>
  ggplot(mapping = aes(x = distance_abs))+
  geom_histogram()

##What causes a big drop in the data? Are there any changes that are bigger than others? 
ideal_aid |> group_by(ccode, session)|>
  summarize(mean_prev_years = sum(idealpin ))
  mutate(change_diff =  )


                    
ideal_aid<- ideal_aid |> 
  select(!session_minus_one)



## W/O USA
ideal_wo_us <- ideal|> 
  group_by(country_cown, year_match)|>
  select(ccode, session, IdealPointAll, Countryname, year_match, country_cown, country_cowc)|>
  filter(!country_cown == 2)

## US Values only
ideal_onlyus <- ideal|>
  filter(country_cown == 2)|>
  select(session, IdealPointAll)|>
  mutate(idealpoint_us = IdealPointAll)|>
  select(!IdealPointAll)

## ideal now has US value
ideal<- ideal|> 
  left_join(ideal_onlyus, by = join_by(session))

ideal <- ideal|> 
  mutate(distance = idealpoint_us - IdealPointAll)|>
  mutate(distance_abs = abs(distance))

View(ideal_us)




```



```{r}

##Using USD for some reggresions 
ideal_aid|>
  ggplot(mapping = aes(x =  committed_millions,
                       y = distance_abs))+
  geom_point()

fit <- lm(distance_abs ~ committed_millions, data = ideal_aid)

summary(fit)

View(fit)
coef(fit)


##Using USD
ideal_treat|> 
  ggplot(mapping = aes(x = committed_millions,
                       y = change_one_before))+
  geom_point()

fit_usd_one <-  lm(change_one_before ~ committed_millions, data = ideal_treat)
fit_usd_two <-  lm(change_two_before ~ committed_millions, data = ideal_treat)

```




```{r loadata, echo = FALSE}
library(tidyverse)
library(dplyr)
install.packages("lubridate")
install.packages("countrycode")
library(lubridate)
library(countrycode)

load("~/Desktop/GOV 50/Projects/Final/final_project/data/un_votes.RData") ## 

ideal <-  read.csv("data/IdealpointestimatesAll_Jul2023.csv")
lead <- read_dta("data/WhyLeadersFightLEADDataset_updated.dta")
presidents <- read.csv("data/presidents.csv")

presidents <- presidents|>
  select(Presidency, President, Took.office, Left.office, Party)

View(ideal)
View(completeVotes)
View(presidents)
## Create the presidents tibble for dates. 

completeVotes|> summary()

```










```{r editingideal}
### Adding new colums for other years
ideal_aid <- ideal_aid|> left_join(minus_one, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_minus_one"))

ideal_aid <- ideal_aid|> left_join(minus_two, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_minus_two"))

ideal_aid <- ideal_aid|> left_join(plus_one, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_plus_one"))


ideal_aid <- ideal_aid|> left_join(plus_two, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_plus_two"))

## Ordering and new titles for columns 
ideal_aid<- ideal_aid|> 
  mutate(distance_two_before = distance_abs_plus_two,
         distance_one_before = distance_abs_plus_one,
         distance_duplicate = distance_abs,
         distance_one_after = distance_abs_minus_one,
         distance_two_after = distance_abs_minus_two)

## Mutating for difference 
ideal_aid<- ideal_aid|> 
  mutate(change_one_before = distance_one_before - distance_duplicate,
         change_two_before = distance_two_before - distance_duplicate)
```


```{r fromresults}
fit_treat<- lm(change_one_before ~ treatment_any, data = ideal_treat)
summary(fit_treat)

## Making distribution of the differences for all sessions 
ate_table|> ggplot( mapping = aes(x = `0`))+
  geom_histogram()

ate_table|> ggplot( mapping = aes(x = `1`))+
  geom_histogram()

## Making point chart for means of sessions
ate_table|> ggplot(mapping = aes( x = ))



##Calculating Means of all sessions 
ate_averages_across <- c(mean(ate_table_names$Treatment), mean(ate_table_names$Control))
as_tibble(ate_averages_across)|> 

ate_averages_across<- data.frame(Treatment = mean(ate_table_names$Treatment),
                                 Control = mean(ate_table_names$Control)
                                 )
  
ate_average_across_table<- as_tibble(ate_averages_across)|> 
  mutate(ate_across_sessions =  `Treatment` - `Control`)

colnames(ate_average_across_table)<- c("Treatment Mean Sessions", "Control Mean Across Sessions", "ATE Across Sessions")

```


```{r}


ideal_treat<- ideal_aid|> mutate(treatment_any =  if_else(committed_millions > 0, 1, 0),
                                 treat_some = if_else( (committed_millions > 0) & (committed_millions <= 1), 1, 0),
                                 treat_middle = if_else( (committed_millions > 1) & (committed_millions <= 100),1,0),
                                 treat_high = if_else(committed_millions >100, 1, 0),
                                 treat_sig = if_else(committed_millions > 10, 1, 0)
                                )

## Histogram of Changes
 ideal_aid|> 
  ggplot(mapping = aes(x = change_one_before))+
  geom_histogram()

 ideal_aid|> 
  ggplot(mapping = aes(x = change_two_before))+
  geom_histogram()
 






```




```{r}
 

minus_one <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, distance_abs)|>
  mutate(session = session - 1)

minus_two <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, distance_abs)|>
  mutate(session = session - 2)
 
plus_one <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, distance_abs)|>
  mutate(session = session + 1 )

plus_two <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, distance_abs)|>
  mutate(session = session + 2)

### Adding new colums for other years
ideal_dis <- ideal_aid|> left_join(minus_one, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_minus_one"))

ideal_aid <- ideal_aid|> left_join(minus_two, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_minus_two"))

ideal_aid <- ideal_aid|> left_join(plus_one, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_plus_one"))


ideal_aid <- ideal_aid|> left_join(plus_two, 
                      by = join_by(ccode, session),
                      suffix =  c("", "_plus_two"))

## Ordering and new titles for columns 
ideal_aid<- ideal_aid|> 
  mutate(distance_two_before = distance_abs_plus_two,
         distance_one_before = distance_abs_plus_one,
         distance_duplicate = distance_abs,
         distance_one_after = distance_abs_minus_one,
         distance_two_after = distance_abs_minus_two)

## Mutating for difference 
ideal_aid<- ideal_aid|> 
  mutate(change_one_before = distance_one_before - distance_duplicate,
         change_two_before = distance_two_before - distance_duplicate)


```



## Finding prev years into rows
minus_one <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, IdealPointAll)|>
  mutate(session = session - 1)

minus_two <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, IdealPointAll)|>
  mutate(session = session - 2)
 
plus_one <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, IdealPointAll)|>
  mutate(session = session + 1 )

plus_two <- ideal_aid|> 
  group_by(ccode, session)|>
  select(ccode, session, IdealPointAll)|>
  mutate(session = session + 2)


```{r}

##Simple Population Means
ideal_treat|> group_by(treatment_any)|> 
  summarize(total_change = mean(change_one_before, na.rm = TRUE))

## Getting just country data for one session, example.
ideal_treat|> 
  filter(session == 29)|>
  select(session, Countryname, committed_millions, treatment_any, IdealPointAll, idealpoint_us, distance_abs, change_one_before)

## Treatment Means and ATE per session, for any aid
test<- ideal_treat|> group_by(session, treatment_any)|>
  summarize(mean_change_one = mean(change_one_before, na.rm = TRUE))|>
  pivot_wider(names_from = treatment_any,
              values_from = mean_change_one)|>
  mutate(session_ate = `1` - `0`)|>
  filter(!is.na(session_ate))

 ## Mean across sessions, any aid 
mean(test$session_ate)

##Regression with treatment, any aid 
fit_treat<- lm(change_one_before ~ treatment_any, data = ideal_treat)
summary(fit_treat)$ coefficients








## Simple means for sig aid
ideal_treat|> group_by(treat_sig)|> 
  summarize(total_change = mean(change_one_before, na.rm = TRUE))

## Treatment Means and ATE per session for significant aid
test_2<- ideal_treat|> group_by(session, treat_sig)|>
  summarize(mean_change_one = mean(change_one_before, na.rm = TRUE))|>
  pivot_wider(names_from = treat_sig,
              values_from = mean_change_one)|>
  mutate(session_ate = `1` - `0`)|>
  filter(!is.na(session_ate))

## Mean across sessions, sig aid
mean(test_2$session_ate)
```

```{r dataglance, eval = FALSE}

## Familiarizing Data.
aid_us|> ggplot(mapping = aes( x = year, y = commitment_amount_usd_constant_sum ))+
  geom_point()

country_amount <-aid_us |> 
   group_by(year_match)|> 
   summarize(num_countries  = length(unique(country_cown)),
             total_given = sum(commitment_amount_usd_constant_sum)) 
  
aid_us |>
  ggplot(mapping = aes(x = as.factor(year_match),
                               y = commitment_amount_usd_constant_sum) )+
  geom_boxplot()+
  scale_y_log10()

country_amount|>
  ggplot(mapping = aes(x = as.factor(year_match),
                       y = num_countries))+
  geom_line()


country_amount|>
  ggplot(mapping = aes(x = year_match,
                       y = total_given))+
  geom_line()
```

### Further Analysis 
```{r}
## Density Chart for ATE Table
ate_table_longer|> 
  mutate(treatment = as.factor(treatment))|>
  ggplot( mapping = aes(x = change_one_before,
                        fill = treatment,
                        color =  treatment))+
  geom_density(alpha = 0.3)


##Distribution Charts for ATE Table 
ate_table_longer|> 
  mutate(treatment = as.factor(treatment))|>
  ggplot( mapping = aes(x = change_one_before,
                        fill = treatment,
                        color =  treatment))+
  geom_histogram()



##Whisker Charts
ggplot(data = midwest,
       mapping = aes(x = percbelowpoverty,
                     fill = state, color = state)) +
  geom_density(alpha = 0.3)

```